---
title: "Assignment 3: Statistical Findings Memo"
author: "Alex Trepka"
format: 
  pdf:
    documentclass: article
    fontfamily: Inter
    fontsize: 11pt
    margin-left: 1in
    margin-right: 1in
    margin-top: 1in
    margin-bottom: 1in
editor: visual
---

# Assignment 3: "The Leap"

## Conceptual Focus

This assignment covers **Chapter 14: Making Inferences**. In Week 13, we *described* our sample (e.g., "In our sample, Group A used more platforms"). Now, we must *verify* these findings. As your slides state, this is the "leap of faith" from our sample to the population.

Our core question is: "Is the difference I see in my sample 'real'... or could it just be a fluke due to random chance?" We will use **Hypothesis Testing** (and p-values) to answer this, and we'll use the correct test for each job based on your **PowerPoint (Slide 23)**.

------------------------------------------------------------------------

# Statistical Findings Memo

**To:** Alex Chen (Director of Insights)

**From:** Alex Trepka (Head Leaper)

**Subject:** Statistical Verification for 2025 Digital Landscape Brief

This memo formally tests the key patterns identified in our descriptive analysis. The goal is to verify that these findings are statistically significant and not just a fluke of our sample before we present them to the client. All tests use an alpha level of $\alpha = .05$.

## RQ 1: Association (TikTok & Politics)

**Research Question:** Is there a significant *association* between political party (`party_simple`) and using TikTok (`uses_tiktok`)?

**Hypotheses:** \* **H0 (Null):** There is no association between `party_simple` and `uses_tiktok` in the population. \* **HA (Alternative):** There is an association between `party_simple` and `uses_tiktok` in the population.

**Results:** \[The `chisq.test` output will appear below.\]

```{r}
#| echo: false
#| ref.label: "chunk-test-chisq"
# This chunk displays the Chi-Square test output
```

**Decision:** \[Based on the p-value (which is...?), do you "Reject H0" or "Fail to Reject H0"?\]

**Conclusion:** \[Write 1-2 sentences in plain English for the client. Is there a link between politics and TikTok use? What does the data show?\]

------------------------------------------------------------------------

## RQ 2: Difference (Device Type)

**Research Question:** Is there a significant *difference* in social media adoption (`platform_count`) between respondents who took the survey on a PC versus a Smartphone?

**Hypotheses:** \* **H0 (Null):** The mean `platform_count` is the same for PC and Smartphone users in the population. \* **HA (Alternative):** The mean `platform_count` is different for PC and Smartphone users in the population.

**Results:** \[The `t.test` output will appear below.\]

```{r}
#| echo: false
#| ref.label: "chunk-test-ttest"
# This chunk displays the t-test output
```

**Decision:** \[Based on the p-value, do you "Reject H0" or "Fail to Reject H0"?\]

**Conclusion:** \[Write 1-2 sentences for the client. Does device type seem to relate to how many platforms a person uses? What are the mean scores for each group (see `mean in group 1` and `mean in group 2`)?\]

------------------------------------------------------------------------

## RQ 3: Difference (Party & Platform Count)

**Research Question:** Is there a significant *difference* in the mean number of platforms used (`platform_count`) across our three `party_simple` groups?

**Hypotheses:** \* **H0 (Null):** The mean `platform_count` is the same for all three `party_simple` groups in the population. \* **HA (Alternative):** At least one group's mean `platform_count` is different from the others.

**Results (ANOVA):** \[The `aov` output will appear below.\]

```{r}
#| echo: false
#| ref.label: "chunk-test-anova"
# This chunk displays the ANOVA test output
```

**Results (Post-Hoc Test):** \[The `TukeyHSD` output will appear below.\]

```{r}
#| echo: false
#| ref.label: "chunk-test-tukey"
# This chunk displays the TukeyHSD test output
```

**Decision:** \[Was the overall ANOVA significant (look at the `Pr(>F)` value)? Based on the TukeyHSD `p adj` values, which *specific pairs* are significantly different (i.e., have p adj \< .05)?\]

**Conclusion:** \[Write 1-2 sentences for the client. This is the "Dumb ANOVA" and "Smart Post-Hoc" from your slides. Does political affiliation relate to the number of platforms people use? If so, who uses more or fewer platforms?\]

------------------------------------------------------------------------

## Final Discussion

\[Look at your results from all three tests. What is the "big picture" story you are starting to see? What is the most *meaningful* or *important* finding **you would recommend we lead with in the client presentation**, and why?\]

\[**Pedagogical Prompt:** In your answer, explicitly consider the concept of "Significance vs. Meaningfulness" from your PowerPoint (Slide 22). Is your most *statistically significant* finding (lowest p-value) also your most *practically important* finding for the client?\]

\newpage

# Appendix: R Code and Commentary

## 1. Setup

\[Add your commentary.\]

```{r}
#| label: "chunk-setup"
#| echo: true
#| message: false
#| warning: false

library(tidyverse)
load("w144_wrangled.RData") # Load data from Assignment 1
```

## 2. RQ 1: Chi-Square Test

\[Add your commentary. **Why is `chisq.test` the right test for RQ 1?** (Hint: Look at your PowerPoint, Slide 23. What are the *variable types* for `party_simple` and `uses_tiktok`?)\]

```{r}
#| label: "chunk-test-chisq"
#| echo: true

# Create a contingency table
# Skeleton: Use the table() function on your two categorical variables
tiktok_party_table <- table(/* variable1 */, /* variable2 */)

# Run the test
# Skeleton: Use the chisq.test() function on your table
chisq.test(tiktok_party_table)
```

## 3. RQ 2: Independent Samples t-test

\[Add your commentary. **Why is `t.test` the right test for RQ 2?** (Hint: Slide 23. What are the variable types? How many groups are we comparing?)\]

```{r}
#| label: "chunk-test-ttest"
#| echo: true

# Filter the data to include only PC (1) and Smartphone (2) users
device_data <- w144_wrangled %>%
  filter(device_type_w144 == 1 | device_type_w144 == 2)

# Run the t-test
# Skeleton: Use t.test(numeric_var ~ categorical_var, data = ...)
t.test(/* numeric_var */ ~ /* group_var */, data = device_data)
```

## 4. RQ 3: ANOVA & TukeyHSD

\[Add your commentary. **Why is `aov` the right test for RQ 3?** (Hint: Slide 23. How is this different from RQ 2?) What is the `TukeyHSD` test for?\]

```{r}
#| label: "chunk-test-anova"
#| echo: true
#| results: "hide" 
# Hide results here so they can be shown by the next chunk

# Run the ANOVA
# Skeleton: Use aov(numeric_var ~ categorical_var, data = ...)
anova_model <- aov(/* numeric_var */ ~ /* group_var */, data = w144_wrangled)

# Get the summary table
summary(anova_model)
```

```{r}
#| label: "chunk-test-tukey"
#| echo: true

# Run the Post-Hoc Test
# Skeleton: Use TukeyHSD() on your anova_model
TukeyHSD(anova_model)
```
